A Horrocks-type theorem for even orthogonal $K_2$

### 2019-09-05 dl-paper.tex
Some further revisions

\begin{document}
\begin{abstract} We prove a $\mathbb{P}^1$-glueing theorem for even-dimensional orthogonal Steinberg groups. This result, which is an analogue of Horrocks theorem, transfers to the orthogonal case an earlier result of M.~Tulenbaev and is also an analogue of an earlier result of A.~Suslin and V.~Kopeiko. \end{abstract}
\maketitle
\section{Introduction}
The solution of the famous Serre problem on projective modules by D.~Quillen and A.~Suslin was one of the brilliant achievements of mathematics of the 70's, which bolstered the development of algebraic K-theory, see~\cite{Su76, Qu76}. In its weakest form Serre problem asks whether any projective module over a polynomial ring $R = k[x_1,\ldots, x_n]$ over a field $k$ is free. We refer the reader to the textbook~\cite{Lam10} for a comprehensive account on the problem, its history and the subsequent solution.

@@

Recall that to every irreducible root system $\Phi$ and a commutative ring $R$ one can associate two groups: the group $G(\Phi, R)$ called {\it simply-connected Chevalley group} (see~\cite[\S~3]{St71} or~\cite{VP}) and the group $\St(\Phi, R)$ called {\it Steinberg group} (see~\cref{sec:Steinberg-intro} for the definition). There is a well-defined map $\pi \colon \St(\Phi, R) \to G(\Phi, R)$ sending each generator $x_\alpha(\xi)$ to the elementary root unipotent $t_\alpha(\xi)$. The cokernel and the kernel of this map are, by definition, the $\K_1$ and $\K_2$-groups modeled on Chevalley group $G(\Phi, -)$ (see~\cite{St78}).
\[ \xymatrix{ 1 \ar[r] & \K_2(\Phi, R) \ar[r] & \St(\Phi, R) \ar^{\pi}[r] & G(\Phi, R) \ar[r] & \K_1(\Phi, R) \ar[r] & 1.} \]

@@

In~\cite{LS17} the authors have shown that Steinberg groups $\St(\Phi, R)$ satisfy Quillen--Suslin local-global principle provided $\Phi$ is simply-laced and has rank $\geq 3$. This result is one of the ingredients needed in the proof of the $\K_2$-analogue of Serre problem. The aim of the present article is to make yet another step towards the solution of the problem, namely to prove an analogue of Horrocks theorem~\cite{Ho64} for Steinberg groups.

Our main result is the following theorem, which is the orthogonal analogue of~\cite[Theorem~5.1]{Tu83} and the $\K_2$-analogue of~\cite[Theorem~6.8]{Su82} (cf. also with~\cite[Theorem~VI.5.2]{Lam10} and~\cite[Theorem~1.1]{St-poly}).
\begin{theorem}[Horrocks theorem for orthogonal $\K_2$]\label{thm:main} Let $A$ be a commutative ring in which $2$ is invertible. Then for any $\ell \geq 7$ the following commutative square is a pullback square in which all maps are injective:
\begin{equation} \nonumber
  \xymatrix{ \KO_2(2\ell, A) \ar[r] \ar[d] & \KO_2(2\ell, A[X]) \ar[d] \\ \KO_2(2\ell, A[X\inv]) \ar[r] & \KO_2(2\ell, A[X, X\inv]).}  
\end{equation} 
Moreover, the same assertion holds if one replaces the functor $\KO_2(2\ell, -)$ with $\K_2(\rD_\ell, -)$ or $\St(\rD_\ell, -)$. \end{theorem}
In the above statement $\KO_2(2\ell, -)$ denotes the unstable orthogonal $\K_2$-functor (see~\cref{sec:quillen}).

Our proof of~\cref{thm:main} goes as follows. We notice that it suffices to prove the $\St(\rD_\ell, -)$-variant of the theorem, moreover, the proof of the injectivity of
$j_+ \colon \St(\rD_\ell, A[X]) \to \St(\rD_\ell, A[X, X\inv])$ turns out to be the hardest part. After invoking the local-global principle~\cite[Theorem~2]{LS17} the proof reduces to the special case when $A$ is local.
Now if $M$ is the maximal ideal of $A$, the proof of the injectivity of $j_+$ comes down to proving injectivity of the following two maps:
\[ \xymatrix{ \St(\rD_\ell, A[X]) \ar[r] & \St(\rD_\ell, A[X\inv] + M[X]) \ar[r] & \St(\rD_\ell, A[X, X\inv]). }\]
The injectivity of the second map is obtained in~\cref{thm41}. This is the only place in our proof which invokes the assumption that $2$ is invertible. The proof of~\cref{thm41} depends on the stability theorem for higher orthogonal K-groups and also on certain basic computations with Grothendieck--Witt groups (these groups include the stable orthogonal K-groups as a special case, see~\eqref{GW-concrete}). Finally, the injectivity of the first map is obtained in~\cref{thm:P1glueing}, which is a direct generalization of~\cite[Proposition~4.3]{Tu83}. This part of the proof is obtained in somewhat greater generality and is applicable to all simply-laced $\Phi$ of rank $\geq 5$.
%$\rD_\ell$, $\ell=5,6$ and also $\rE_\ell$, $\ell=6,7,8$.

@@

 Set $\mathcal{R}^\Phi_{n+1} = \mathcal{R}^\Phi_{\leq n+1} \setminus \mathcal{R}^\Phi_{\leq n}$. It suffices to verify that the image of every relator $R \in \mathcal{R}^\Phi_{n+1}$ under $\widetilde{j}^\Phi_n$ is a trivial element of $\St^{\leq n}(\Phi, B)$. In the special case $\Phi=\rA_{\geq 4}$ this has already been demonstrated by Tulenbayev in~\cite[Lemma~3.3]{Tu83}, so in this case the proof of the proposition is complete. We will deduce the assertion in the remaining cases $\Phi=\rD_\ell,\rE_\ell$ from the special case $\Phi=\rA_4$ of Tulenbayev's result.

 Let $R$ be a relation from $\mathcal{R}^\Phi_{n+1}$. By~\cref{superfluous-relations} we may assume that $R$ is not of type $\text{R3}^\bot$, hence the roots $\alpha, \beta$ appearing in $R$ are contained in a root subsystem of $\Phi$ of type $\rA_2$. Our assumptions on $\Phi$ guarantee that there exists some root subsystem $\Psi$ of type $\rA_4$ containing $\alpha$ and $\beta$. Consider the following commutative diagram in which the vertical arrows are induced by the embedding $\Psi\subseteq\Phi$. 
  \begin{equation} \nonumber \xymatrix{
 F(\mathcal{X}_{\leq n+1}^\Psi) \ar[r]^{j_n^\Psi} \ar[d] & \St^{\leq n}(\Psi, B) \ar[d] \\
 F(\mathcal{X}_{\leq n+1}^\Phi) \ar[r]^{j_n^\Phi} & \St^{\leq n}(\Phi, B) }
 \end{equation}
The relation $R$ lies in the image of the left arrow, therefore it comes from some relation $R' \in \mathcal{R}^\Psi_{n+1}$. The image of $R'$ in $\St^{\leq n}(\Psi, B)$ under $j_n^\Psi$ is trivial by Tulenbayev's result. But this implies that the image of $R$ under $\widetilde{j}_n^\Phi$ is also trivial and hence that $\widetilde{j}_n^\Phi$ gives rise to the desired map $j_n$.
\end{proof}

@@

\begin{lemma} \label{lem:Tulenbayev-formula} 
One has $S_\alpha(a, x_{-\alpha}(mX)) = x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot \langle a, m\rangle_\alpha \cdot h_\alpha(1+am).$
\end{lemma}
\begin{proof} Since $\Phi$ is nonsymplectic, we can choose $\gamma \in \Phi$ such that $\langle \alpha, \gamma \rangle = -1$. Direct computation using~\eqref{eq:conj-h-x},\eqref{eq:conj-h-h},\eqref{eq:steinberg},\eqref{eq:dennis-stein},\eqref{eq:symbol-properties} shows that
 \begin{multline*}
 x_\alpha(aX^{-1}) \cdot x_{-\alpha}(mX) = {}^{h_\gamma(X)}(x_\alpha(a) \cdot x_{-\alpha}(m)) = \\
 = {}^{h_\gamma(X)}\left( x_{-\alpha}\left(\tfrac{m}{1+am}\right) \cdot \langle a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_\alpha\left(\tfrac{a}{1+am}\right) \right) = \\
 = x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot \langle a, m\rangle_\alpha \cdot h_\alpha(X^{-1}(1+am))\cdot h_\alpha^{-1}\left(X^{-1}\right) \cdot x_{\alpha}\left(\tfrac{aX^{-1}}{1+am}\right) = \\
 = x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot \langle a, m\rangle_\alpha \cdot \{X^{-1}, 1+am\} \cdot h_\alpha(1+am)\cdot x_{\alpha}\left(\tfrac{aX^{-1}}{1+am}\right) = \\
 = x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot \langle a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_{\alpha}\left(\tfrac{aX^{-1}}{1+am}\right) \cdot \{1+am, X\}, \end{multline*}
which implies the assertion. \end{proof}

@@

\begin{lemma} \label{SR:sharp} For $g \in K(A[X], M[X]) \cap P_\alpha(m) \cap P_\beta(m')$ and $a \in A$ holds \[S_\alpha(a, g)\cdot  h^{-1}_\alpha(1 + am) \cdot x_{\alpha-\beta}(-N_{\alpha,-\beta}\cdot am') \in P_\beta\left(\tfrac{m'}{1 + am}\right). \] \end{lemma}
\begin{proof} By~\cref{K-a-b} $g$ can be presented as $ g_0 \cdot x_{-\alpha}(mX) \cdot x_{-\beta}(m'X)$ for some $g_0\in K(\alpha, \beta)$, therefore from~\eqref{eq:Smult} and~\cref{lem:Tulenbayev-formula} we obtain that \begin{multline} \nonumber S_\alpha(a, g) = S_\alpha(a, g_0) \cdot S_\alpha(a, x_{-\alpha}(mX)) \cdot S_\alpha\left(\tfrac{a}{1+am}, x_{-\beta}(m'X)\right) = \\ = S_\alpha(a, g_0) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot \langle a, m \rangle \cdot h_\alpha(1+am) \cdot x_{-\beta}(m'X) \cdot x_{\alpha-\beta}\left(\tfrac{N_{\alpha, -\beta} \cdot am'}{1+am}\right) = \\ = S_\alpha(a, g_0) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot \langle a, m \rangle \cdot  x_{-\beta}\left(\tfrac{m'X}{1+am}\right) \cdot x_{\alpha-\beta}\left(N_{\alpha, -\beta} \cdot am'\right) \cdot h_\alpha(1+am). \end{multline}
The assertion now follows from the preceding corollary.
\end{proof}

@@

For shortness we denote the subgroup $\overline{\St}(\Phi, A, M)$ by $G_M^0$.
Notice that $G_M^0$ can be considered as a subgroup of $\St(\Phi, R)$.
\[ \xymatrix{ G_M^0 \ar@{^{(}->}[r]^{i_+} \ar@{^{(}->}[d]^{i_-} & \overline{\St}(\Phi, A[X], M[X]) \ar^{j_+^M}[d] \ar@{^{(}->}[r] & \St(\Phi, A[X]) \ar[dl]^{j_+} \\
              \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, R) &  } \] 
We denote by $\overline{G}^{\geq 0}_M$ the image of the map $j_+^M$. 
Denote by $V$ the subset of \[\St(\Phi, R) \times \St(\Phi, A[X\inv]) \times (1 + M)^\times\] consisting of those triples $(g, h, u)$ for which
\[p(g, h, u) := g \cdot j_-(h) \cdot \{ X, u \}\] belongs to $\overline{G}_M^{\geq 0}$.

We let $h_0 \in G_M^0$ act on $V$ on the right by $(g, h, u) \cdot h_0 = (g, h \cdot i_-(h_0), u)$. This action is well defined since $G^0_M \subset \overline{G}^{\geq 0}_M$. Whenever two triples $v_1, v_2 \in V$ lie in the same orbit we use the notation $v_1 \sim v_2$ and say that they are {\it congruent}. The same notation $h_1 \sim h_2$ is also used for the elements of $\St(\Phi, A[X\inv])$ lying in the same left $G^0_M$-coset.

We denote by $\overline{V}$ the set of orbits of the action of $G_M^0$.
We use the notation $(g, [h], u)$ for the elements of $\overline{V}$.
We denote by $\overline{p}$ the function $\overline{V} \to \overline{G}^{\geq 0}_M/G_M^0$ sending each $(g, [h], u) \in V$ to the left coset $p(g, h, u)G_M^0$.
The set $\overline{V}$ is the one upon which we will construct an action of $\St(\Phi, B)$.

Before we proceed further let us explain how our sets $V$ and $\overline{V}$ compare with the constructions used by Tulenbaev in his proof of~\cite[Proposition~4.3]{Tu83}. This may also help to explain the intuition behind our definition of $\overline{V}$. Recall that Tulenbaev constructs an action of $\St(\Phi, B)$ (of course, only for $\Phi = \rA_\ell$) on a different set, which we denote by $\overline{V}_T$. This set $\overline{V}_T$ is defined as the quotient of the set of triples $V_T = \overline{G}_M^{\geq 0} \times \St(\Phi, A[X\inv]) \times (1+M)^\times$ modulo the equivalence relation given by $(p \gamma, h, u) \sim (p, \gamma h, u)$, $\gamma \in G^0_M$. Tulenbaev denotes the image of $(p, h, u)$ in $\overline{V}_T$ by $[p, h, u]$.

The isomorphism between the sets $\overline{V}$ and $\overline{V}_T$ is given by the following two maps, which are easily seen to be mutually inverse to each other:
\begin{equation} \label{eq:VVT} \xymatrix @R0.3pc {\overline{V} \ar[r]^{\cong} & \ar[l] \overline{V}_T \\ (g, [h], u) \ar@{|->}[r] & [p(g, h, u), h^{-1}, u] \\ (p \cdot j_-(h) \cdot \{u, X\}, [h^{-1}], u) & \ar@{|->}[l] [p, h, u]. } \end{equation}
The above isomorphism allows us to treat $\overline{V}$ and $\overline{V}_T$ like the same object, for which we can interchangeably use either of the two notations,
 depending on which of them is more convenient in a given situation.
For example, specifying the action of $\St(\Phi, B)$ in terms of $\overline{V}$ leads to much shorter calculations in Lemmas~\ref{R3_leq0_1}--\ref{R2_0_1},
 while the statements of~\cref{prop43} and~\cref{lem:action} look more natural when formulated in terms of~$\overline{V}_T$.

@@

\begin{lemma}\label{lem:orbit-action} Let $(A, M)$ be a local pair. 
Then for any $\alpha \in \Phi$ and $a \in A$ the partial function $t_\alpha(aX^{-1}) \colon V \not \to V$ gives rise to a well-defined total function $T_\alpha(aX^{-1}) \colon \overline{V} \to \overline{V}$. \end{lemma}
\begin{proof}
 First of all, let us show that the resulting function is total. 
 Fix $(g, h, u) \in V$. Since $p(g, h, u) \in \overline{G}^{\geq 0}_M$ there exists $g_1 \in \overline{\St}(\Phi, A[X], M[X])$ such that $j_+(g_1) = p(g, h, u).$
 Set $h_0 = ev^*_{X=0}(g_1)^{-1}$, then, clearly, $g_1 \cdot i_+(h_0) \in K(A[X], M[X])$ and 
 $p(g, h \cdot i_-(h_0), u) = j_+(g_1) \cdot j_-(i_-(h_0)) = j_+(g_1 \cdot i_+(h_0)) \in j_+(K(A[X], M[X])) \subseteq P_\alpha(*)$, which shows that $t_\alpha(aX^{-1})$ is defined on the representative $(g, h \cdot i_+(h_0), u)$, which is congruent to $(g, h, u)$.

 Next, let us show that the value of $T_\alpha(aX^{-1})$ does not depend on a choice of the representative.
 Let $(g, h_1, u) \sim (g, h_2, u)$ be elements of the same orbit for which both $p(g, h_1, u)$ and $p(g, h_2, u)$ belong to $P_\alpha(*)$.
 By definition, $h_1^{-1} h_2 = i_-(h_0)$, for some $h_0 \in G^0_M$, moreover, 
  $p(g, h_1, u)^{-1} \cdot p(g, h_2, u) = j_- i_-(h_0) = j_+ i_+(h_0) \in P_\alpha(*)$.
 By~\cref{Pstar-char} there exists $g_1 \in \overline{\St}(\Phi, A[X], M[X])$ such that $g_0 := ev^*_{X=0}(g_1) \in Z_\alpha(A, M)$ and $j_+(g_1) = j_+ i_+ (h_0)$.
 From the last equality and the injectivity of the map $G(\Phi, A[X]) \to G(\Phi, R)$ we obtain that the projections of $g_0$, $g_1$ and $h_0$ in $G(\Phi, R)$ are equal, which shows that $g_0 \cdot h_0^{-1} \in \overline{\K_2}(\Phi, A, M)$. It follows from~\cref{thm:Stein} that the latter subgroup is generated by relative Steinberg symbols $\{ a, 1 + m \}$ and hence by~\cref{Z-DS} it is contained in $Z_\alpha(A, M)$. Thus, we have obtained that $h_0 \in Z_\alpha(A, M)$ and hence that $i_-(h_0)$ is centralized by $X_\alpha(A[X^{-1}])$, which allows us to conclude that
 \begin{multline} \nonumber
  t_\alpha(aX^{-1})(g, h_1, u) = \left( x_\alpha(aX^{-1})\cdot g ,\ h_1 \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ u \cdot (1 + am)\right) \sim \\
  \sim \left( x_\alpha(aX^{-1})\cdot g ,\ h_1 i_-(h_0) \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ u \cdot (1 + am)\right) = t_\alpha(aX^{-1})(g, h_2, u). \qedhere
 \end{multline} \end{proof}

>>>>

\begin{document}
\begin{abstract} We prove a $\mathbb{P}^1$-glueing theorem for even-dimensional orthogonal Steinberg groups. This result transfers to the orthogonal case an earlier result of M.~Tulenbaev and is also an analogue of an earlier result of A.~Suslin and V.~Kopeiko. \end{abstract}
\maketitle
\section{Introduction}
Recall that the classical Serre problem on projective modules asks if any projective module over a polynomial ring $R = k[x_1,\ldots, x_n]$ over a field $k$ is free. The problem was positively settled by D.~Quillen~\cite{Qu76} and A.~Suslin~\cite{Su76}, and the solution of the problem played an important role in the development of algebraic K-theory. We also refer the reader to the textbook~\cite{Lam10} for a comprehensive account on the problem, its history and the subsequent solution.

@@

Recall that to every irreducible root system $\Phi$ and a commutative ring $R$ one can associate two groups: the {\it simply-connected Chevalley group} $G(\Phi, R)$ (see~\cite[\S~3]{St71} or~\cite{VP}) and the {\it Steinberg group} $\St(\Phi, R)$ (see~\cref{sec:Steinberg-intro} for the definition). There is a well-defined map $\pi \colon \St(\Phi, R) \to G(\Phi, R)$ sending each generator $x_\alpha(\xi)$ to the elementary root unipotent $t_\alpha(\xi)$. The cokernel and the kernel of this map are, by definition, the $\K_1(\Phi, R)$ and $\K_2(\Phi, R)$-functors modeled on Chevalley group $G(\Phi, -)$, see~\cite{St78}.

@@

In~\cite{LS17} the authors have shown that Steinberg groups $\St(\Phi, R)$ satisfy Quillen--Suslin local-global principle provided $\Phi$ is simply-laced and has rank $\geq 3$. This result is one of the ingredients needed in the proof of the $\K_2$-analogue of Serre problem. The aim of the present article is to make yet another step towards its solution, namely to prove an analogue of Horrocks theorem~\cite{Ho64} for Steinberg groups.

Our main result is, thus, the following theorem, which is the orthogonal analogue of~\cite[Theorem~5.1]{Tu83} and the $\K_2$-analogue of~\cite[Theorem~6.8]{Su82} (cf. also with~\cite[Theorem~VI.5.2]{Lam10} and~\cite[Theorem~1.1]{St-poly}).
\begin{theorem}[Horrocks theorem for orthogonal $\K_2$]\label{thm:main} Let $A$ be a commutative ring in which $2$ is invertible. Then for any $\ell \geq 7$ the following commutative square is a pullback square in which all maps are injective:
\begin{equation} \nonumber
  \xymatrix{ \KO_2(2\ell, A) \ar[r] \ar[d] & \KO_2(2\ell, A[X]) \ar[d] \\ \KO_2(2\ell, A[X\inv]) \ar[r] & \KO_2(2\ell, A[X, X\inv]).}  
\end{equation} 
Moreover, the same assertion holds if one replaces the functor $\KO_2(2\ell, -)$ with $\K_2(\rD_\ell, -)$ or $\St(\rD_\ell, -)$. \end{theorem}
In the above statement $\KO_2(2\ell, -)$ denotes the unstable orthogonal $\K_2$-functor (see~\cref{sec:quillen}).

The proof of~\cref{thm:main} goes as follows. We notice that it suffices to prove the $\St(\rD_\ell, -)$-variant of the theorem. Moreover, the proof of the injectivity of
$j_+ \colon \St(\rD_\ell, A[X]) \to \St(\rD_\ell, A[X, X\inv])$ turns out to be the hardest part. After invoking the local-global principle~\cite[Theorem~2]{LS17} the proof reduces to the special case when $A$ is local.
Now if $M$ is the maximal ideal of $A$, the proof of the injectivity of $j_+$ comes down to proving injectivity of the following two maps:
\[ \xymatrix{ \St(\rD_\ell, A[X]) \ar[r] & \St(\rD_\ell, A[X\inv] + M[X]) \ar[r] & \St(\rD_\ell, A[X, X\inv]). }\]
The injectivity of the second map is obtained in~\cref{thm41}. This is the only place in our proof which invokes the assumption that $2$ is invertible. The proof of~\cref{thm41} depends on the stability theorem for higher orthogonal K-groups and also on certain basic computations with Grothendieck--Witt groups (these groups include the stable orthogonal K-groups as a special case, see~\eqref{GW-concrete}). Finally, the injectivity of the first map is obtained in~\cref{thm:P1glueing}, which is a direct generalization of~\cite[Proposition~4.3]{Tu83}. This part of the proof is obtained in somewhat greater generality and is applicable to all simply-laced root systems $\Phi$ of rank $\geq 5$.
%$\rD_\ell$, $\ell=5,6$ and also $\rE_\ell$, $\ell=6,7,8$.

@@

 Set $\mathcal{R}^\Phi_{n+1} = \mathcal{R}^\Phi_{\leq n+1} \setminus \mathcal{R}^\Phi_{\leq n}$. It suffices to verify that the image of every relator $R \in \mathcal{R}^\Phi_{n+1}$ under $\widetilde{j}^\Phi_n$ is a trivial element of $\St^{\leq n}(\Phi, B)$. In the special case $\Phi=\rA_{\geq 4}$ this has already been demonstrated by Tulenbaev in~\cite[Lemma~3.3]{Tu83}, so in this case the proof of the proposition is complete. We will deduce the assertion in the remaining cases $\Phi=\rD_\ell,\rE_\ell$ from the special case $\Phi=\rA_4$ of Tulenbaev's result.

 Let $R$ be a relation from $\mathcal{R}^\Phi_{n+1}$. By~\cref{superfluous-relations} we may assume that $R$ is not of type $\text{R3}^\bot$, hence the roots $\alpha, \beta$ appearing in $R$ are contained in a root subsystem of $\Phi$ of type $\rA_2$. Our assumptions on $\Phi$ guarantee that there exists some root subsystem $\Psi$ of type $\rA_4$ containing $\alpha$ and $\beta$. Consider the following commutative diagram in which the vertical arrows are induced by the embedding $\Psi\subseteq\Phi$. 
  \begin{equation} \nonumber \xymatrix{
 F(\mathcal{X}_{\leq n+1}^\Psi) \ar[r]^{j_n^\Psi} \ar[d] & \St^{\leq n}(\Psi, B) \ar[d] \\
 F(\mathcal{X}_{\leq n+1}^\Phi) \ar[r]^{j_n^\Phi} & \St^{\leq n}(\Phi, B) }
 \end{equation}
The relation $R$ lies in the image of the left arrow, therefore it comes from some relation $R' \in \mathcal{R}^\Psi_{n+1}$. The image of $R'$ in $\St^{\leq n}(\Psi, B)$ under $j_n^\Psi$ is trivial by Tulenbaev's result. But this implies that the image of $R$ under $\widetilde{j}_n^\Phi$ is also trivial and hence that $\widetilde{j}_n^\Phi$ gives rise to the desired map $j_n$.
\end{proof}

@@

\begin{lemma} \label{lem:Tulenbaev-formula} 
One has $S_\alpha(a, x_{-\alpha}(mX)) = x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot \langle a, m\rangle_\alpha \cdot h_\alpha(1+am).$
\end{lemma}
\begin{proof} Since $\Phi$ is nonsymplectic, we can choose $\gamma \in \Phi$ such that $\langle \alpha, \gamma \rangle = -1$. Direct computation using~\eqref{eq:conj-h-x},\eqref{eq:conj-h-h},\eqref{eq:steinberg},\eqref{eq:dennis-stein},\eqref{eq:symbol-properties} shows that
 \begin{multline*}
 x_\alpha(aX^{-1}) \cdot x_{-\alpha}(mX) = {}^{h_\gamma(X)}(x_\alpha(a) \cdot x_{-\alpha}(m)) = \\
 = {}^{h_\gamma(X)}\left( x_{-\alpha}\left(\tfrac{m}{1+am}\right) \cdot \langle a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_\alpha\left(\tfrac{a}{1+am}\right) \right) = \\
 = x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot \langle a, m\rangle_\alpha \cdot h_\alpha(X^{-1}(1+am))\cdot h_\alpha^{-1}\left(X^{-1}\right) \cdot x_{\alpha}\left(\tfrac{aX^{-1}}{1+am}\right) = \\
 = x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot \langle a, m\rangle_\alpha \cdot \{X^{-1}, 1+am\} \cdot h_\alpha(1+am)\cdot x_{\alpha}\left(\tfrac{aX^{-1}}{1+am}\right) = \\
 = x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot \langle a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_{\alpha}\left(\tfrac{aX^{-1}}{1+am}\right) \cdot \{1+am, X\}, \end{multline*}
which implies the assertion. \end{proof}

@@

\begin{lemma} \label{SR:sharp} For $g \in K(A[X], M[X]) \cap P_\alpha(m) \cap P_\beta(m')$ and $a \in A$ holds \[S_\alpha(a, g)\cdot  h^{-1}_\alpha(1 + am) \cdot x_{\alpha-\beta}(-N_{\alpha,-\beta}\cdot am') \in P_\beta\left(\tfrac{m'}{1 + am}\right). \] \end{lemma}
\begin{proof} By~\cref{K-a-b} $g$ can be presented as $ g_0 \cdot x_{-\alpha}(mX) \cdot x_{-\beta}(m'X)$ for some $g_0\in K(\alpha, \beta)$, therefore from~\eqref{eq:Smult} and~\cref{lem:Tulenbaev-formula} we obtain that \begin{multline} \nonumber S_\alpha(a, g) = S_\alpha(a, g_0) \cdot S_\alpha(a, x_{-\alpha}(mX)) \cdot S_\alpha\left(\tfrac{a}{1+am}, x_{-\beta}(m'X)\right) = \\ = S_\alpha(a, g_0) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot \langle a, m \rangle \cdot h_\alpha(1+am) \cdot x_{-\beta}(m'X) \cdot x_{\alpha-\beta}\left(\tfrac{N_{\alpha, -\beta} \cdot am'}{1+am}\right) = \\ = S_\alpha(a, g_0) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot \langle a, m \rangle \cdot  x_{-\beta}\left(\tfrac{m'X}{1+am}\right) \cdot x_{\alpha-\beta}\left(N_{\alpha, -\beta} \cdot am'\right) \cdot h_\alpha(1+am). \end{multline}
The assertion now follows from the preceding corollary.
\end{proof}

@@

For shortness we denote the subgroup $\overline{\St}(\Phi, A, M)$ by $G_M^0$.
Notice that $G_M^0$ can be considered as a subgroup of $\St(\Phi, R)$.
\[ \xymatrix{ G_M^0 \ar@{^{(}->}[r]^{i_+} \ar@{^{(}->}[d]^{i_-} & \overline{\St}(\Phi, A[X], M[X]) \ar^{j_+^M}[d] \ar@{^{(}->}[r] & \St(\Phi, A[X]) \ar[dl]^{j_+} \\
              \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, R). &  } \] 
We denote by $\overline{G}^{\geq 0}_M$ the image of the map $j_+^M$.               
Denote by $\overline{V}_T$ the quotient of the set of triples 
\[V_T = \overline{G}_M^{\geq 0} \times \St(\Phi, A[X\inv]) \times (1+M)^\times\] by the equivalence relation given by $(p \gamma, h, u)_T \sim (p, \gamma h, u)_T$, $\gamma \in G^0_M$. We denote the image of $(p, h, u)_T$ in $\overline{V}_T$ by $[p, h, u]$.              
$\overline{V}_T$ is precisely the set upon which M.~Tulenbaev constructs an action of $\St(\Phi, B)$ in~\cite[Proposition~4.3]{Tu83}.

Sometimes it will be more convenient for us to work with another set $\overline{V}$ isomorphic to $\overline{V}_T$. Denote by $V$ the subset of $\St(\Phi, R) \times \St(\Phi, A[X\inv]) \times (1 + M)^\times$ consisting of those triples $(g, h, u)$ for which $p(g, h, u) := g \cdot j_-(h) \cdot \{ X, u \}$ belongs to $\overline{G}_M^{\geq 0}$.

We let $h_0 \in G_M^0$ act on $V$ on the right by $(g, h, u) \cdot h_0 = (g, h \cdot i_-(h_0), u)$. This action is well defined since $G^0_M \subset \overline{G}^{\geq 0}_M$.
We denote by $\overline{V}$ the set of orbits of this action of $G_M^0$ and use the notation $(g, [h], u)$ for the elements of $\overline{V}$.
Whenever $v_1, v_2 \in V$ lie in the same $G_M^0$-orbit we use the notation $v_1 \sim v_2$.
We denote by $\overline{p}$ the function $\overline{V} \to \overline{G}^{\geq 0}_M/G_M^0$ sending each $(g, [h], u) \in V$ to the left coset $p(g, h, u)G_M^0$.

The isomorphism between the sets $\overline{V}$ and $\overline{V}_T$ is given by the following two maps, which are easily seen to be mutually inverse to each other:
\begin{equation} \label{eq:VVT} \xymatrix @R0.3pc {\overline{V} \ar[r]^{\cong} & \ar[l] \overline{V}_T \\ (g, [h], u) \ar@{|->}[r] & [p(g, h, u), h^{-1}, u] \\ (p \cdot j_-(h) \cdot \{u, X\}, [h^{-1}], u) & \ar@{|->}[l] [p, h, u]. } \end{equation}
The above isomorphism allows us to regard $\overline{V}$ and $\overline{V}_T$ as one and the same object, for which we can interchangeably use either of the two notations,
 depending on which of them is more convenient in a given situation.
For example, specifying the action of $\St(\Phi, B)$ in terms of $\overline{V}$ leads to much shorter calculations in Lemmas~\ref{R3_leq0_1}--\ref{R2_0_1},
 while the statements of~\cref{prop43} and~\cref{lem:action} look more natural when formulated in terms of~$\overline{V}_T$.

@@

\begin{lemma}\label{lem:orbit-action} Let $(A, M)$ be a local pair. 
Then for any $\alpha \in \Phi$ and $a \in A$ the partial function $t_\alpha(aX^{-1}) \colon V \not \to V$ gives rise to a well-defined total function $T_\alpha(aX^{-1}) \colon \overline{V} \to \overline{V}$. \end{lemma}
\begin{proof}
 First of all, let us show that the resulting function is total. 
 Fix $v_0 = (g, h, u) \in V$. Since $p(g, h, u) \in \overline{G}^{\geq 0}_M$ there exists $g_1 \in \overline{\St}(\Phi, A[X], M[X])$ such that $j_+(g_1) = p(g, h, u).$
 Set $h_0 = ev^*_{X=0}(g_1)^{-1}$, then, clearly, $g_1 \cdot i_+(h_0) \in K(A[X], M[X])$ and 
 $p(g, h \cdot i_-(h_0), u) = j_+(g_1) \cdot j_-(i_-(h_0)) = j_+(g_1 \cdot i_+(h_0)) \in j_+(K(A[X], M[X])) \subseteq P_\alpha(*)$, which shows that $t_\alpha(aX^{-1})$ is defined on the representative $(g, h \cdot i_+(h_0), u)$ lying in the same $G_M^0$-orbit as $v_0$.
